---
title: "HW4"
output: html_document
date: "2023-01-05"
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(tidymodels)
library(schrute)
library(lubridate)
library(knitr)
library(openintro)
library(ROSE)
```





##################################

 
```{r lines, echo=FALSE}
set.seed(1116)
train = read_rds("train_data.csv")
test = read_rds("test_data.csv")
```


```{r}
vars <- train %>% select("attach", "line_breaks", "image", "dollar", "cc")

vars_mean <- sapply(vars, mean)

vars_median <- sapply(vars, median)

vars_std <- sapply(vars, sd)

vars_summary <- tibble(
  Variable = c("attach", "line_breaks", "image", "dollar", "cc"),
  Mean = vars_mean,
  Median = vars_median,
  STD = vars_std
)

vars_summary

```

```{r}
ggplot(train, aes(x=attach)) +
  geom_histogram(binwidth = 0.5,fill="pink",color="black") +
  xlim(-1,5) +
  labs(x="number of attachments", y="frequency") +
  stat_bin(aes(label = ifelse(..count.. == 0, "", paste0(round((..count..)/sum(..count..)*100), "%"))), 
           geom = "text", vjust = -0.5)
```

```{r}
ggplot(train, aes(x=line_breaks)) +
  geom_histogram(fill="green", color= "black") +
  labs(x="number of line breaks", y="frequency") +
  stat_bin(aes(label = ifelse(round((..count..)/sum(..count..)*100) == 0, "",
                              paste0(round((..count..)/sum(..count..)*100),"%"))),geom = "text", vjust = -0.5)
```

```{r}
ggplot(train, aes(x=image)) +
  geom_histogram(binwidth=0.5,fill="blue",color="black") +
  xlim(-1,5) +
  labs(x="number of images", y="frequency") +
  stat_bin(aes(label = ifelse(round((..count..)/sum(..count..)*100) == 0, "",
                              paste0(round((..count..)/sum(..count..)*100),"%"))),geom = "text", vjust = -0.5)
```

```{r}
ggplot(train, aes(x=dollar)) +
  geom_histogram(binwidth=0.8,fill="yellow",color="black") +
  xlim(-1,45) +
  labs(x="number of dollars sign", y="frequency")
```

```{r}
ggplot(train, aes(x=cc)) +
  geom_histogram(color="black",binwidth=0.7,fill="lightblue") +
  xlim(-1,37) 

```


**There are several issues with the data:**

Firstly, the "viagra" variable has only one non-zero value, which is 8, while all other values are 0. This is a problem because it could lead to incorrect results in the model and the variable may not be necessary as a feature.
Moreover, we can see that also the "from" variable has the same issue - it has only two 0 values and the rest values is 1.

Secondly, the "time" variable contains both date and time information together. This makes it difficult to include it in the model and it needs to be separated into two different variables.

Finally, "num_char" and "line_breaks" variables having too many unique values it might make it difficult to detect patterns or perform analyses.


**2. logistic regression**

```{r}
email_rec <- recipe(spam ~ ., data = train) %>%
  step_rm(from, sent_email) %>%
  step_date(time, features = c("dow", "month")) %>%               
  step_rm(time) %>%
  step_cut(cc, attach, dollar, breaks = c(0, 1)) %>%
  step_cut(inherit, password, breaks = c(0, 1, 5, 10, 20)) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_zv(all_predictors())
```

```{r}
email_mod <- logistic_reg() %>% 
  set_engine("glm")
email_mod
```

```{r}
email_wflow <- workflow() %>% 
  add_model(email_mod) %>% 
  add_recipe(email_rec)
```


```{r}
email_fit <- email_wflow %>% 
  fit(data = train)
```

```{r}
tidy(email_fit) %>% print(n = 31)
```

```{r}
email_pred <- predict(email_fit, test) %>% 
  bind_cols(test)

email_pred
```

```{r}
email_pred$spam <- as.factor(email_pred$spam)
conf_matrix <- conf_mat(email_pred, truth = spam, estimate = .pred_class)
summary(conf_matrix)

```

**the result of this model is:**<br>
precision value: 0.9196507<br>
recall value: 0.9887324<br>
F1 value: 0.9529412<br>

To improve this model, I decided to look at the F1 score value which combines both the precision and the recall. In order to decide which features to involve in the model or to omit, I looked at the coefficients of each one of the variables in the email_pred table. I noticed that the coefficients for viagra, line_breaks, exclaim_mess and num_char are low, indicating that they do not have a high effect on the model results. Therefore, I chose to remove them from the model and checked if the F1 score would improve. Additionally, I checked the sent_email column and found that there are many unique values that can help the model identify spam emails, so I decided to add it to the model and check if it improves the performance.


```{r}
email_rec1 <- recipe(spam ~ ., data = train) %>%
  step_rm(from, viagra, line_breaks,exclaim_mess,num_char) %>%
  step_date(time, features = c("dow", "month")) %>%               
  step_rm(time) %>%
  step_cut(cc, attach, dollar, breaks = c(0, 1)) %>%
  step_cut(inherit, password, breaks = c(0, 1, 5, 10, 20)) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_zv(all_predictors())
```

```{r}
email_mod1 <- logistic_reg() %>% 
  set_engine("glm")
email_mod1
```

```{r}
email_wflow1 <- workflow() %>% 
  add_model(email_mod1) %>% 
  add_recipe(email_rec1)
```


```{r}
email_fit1 <- email_wflow1 %>% 
  fit(data = train)
```

```{r}
tidy(email_fit1) %>% print(n = 31)
```

```{r}
email_pred1 <- predict(email_fit1, test) %>% 
  bind_cols(test)

email_pred1
```

```{r}
email_pred1$spam <- as.factor(email_pred1$spam)
conf_matrix1 <- conf_mat(email_pred1, truth = spam, estimate = .pred_class)
summary(conf_matrix1)
```

**The results of the final model is:**<br>
precision value: 0.9308511<br>
recall value: 0.9859155<br>
F1 value: 0.9575923<br>

**After changing the features in the model, I obtained improved results. I ran the model multiple times with different combinations of features that I mentioned before and found that this particular combination of features produces the best F1 score value.**
